---
title: "Data Visualization Exercises"
output: html_notebook
---

```{r}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

# 1.2.5 Exercises

1.  How many rows are in `penguins`? How many columns?

```{r}
dim(penguins)
```

Rows: 344, Columns: 8

2.  What does the `bill_depth_mm` variable in the `penguins` data frame describe? Read the help for `?penguins` to find out.

a number denoting bill depth (millimeters)

3.  Make a scatterplot of `bill_depth_mm` vs. `bill_length_mm`. That is, make a scatterplot with `bill_depth_mm` on the y-axis and `bill_length_mm` on the x-axis. Describe the relationship between these two variables.

```{r}
penguins |> 
  ggplot(aes(y = bill_depth_mm, x = bill_length_mm)) +
  geom_point(na.rm = TRUE)
```

Positive linear correlation

4.  What happens if you make a scatterplot of `species` vs. `bill_depth_mm`? What might be a better choice of geom?

```{r}
penguins |> 
  ggplot(aes(bill_depth_mm, species)) +
  geom_point(na.rm = TRUE)
```

```{r}
penguins |> 
  ggplot(aes(species)) +
  geom_bar()
```

5.  Why does the following give an error and how would you fix it?

    ```{r, eval=FALSE}

    ggplot(data = penguins) + 
      geom_point()
    ```

    Missing aesthetics

6.  What does the `na.rm` argument do in `geom_point()`? What is the default value of the argument? Create a scatterplot where you successfully use this argument set to `TRUE`.

```{r}
penguins |> 
  ggplot(aes(bill_length_mm, bill_depth_mm)) +
  geom_point(na.rm = TRUE)
```

7.  Add the following caption to the plot you made in the previous exercise: "Data come from the palmerpenguins package." Hint: Take a look at the documentation for `labs()`.

    ```{r}
    penguins |> 
      ggplot(aes(bill_depth_mm, bill_length_mm)) +
      geom_point(na.rm = TRUE) +
      labs(caption = "Data come from the palmerpenguins package.")

    ```

8.  Recreate the following visualization. What aesthetic should `bill_depth_mm` be mapped to? And should it be mapped at the global level or at the geom level?

    ```{r}
    ggplot(
      penguins, aes(flipper_length_mm, body_mass_g)) +
      geom_point(aes(color = bill_depth_mm), na.rm = TRUE) +
      geom_smooth(method = "lm")
    ```

9.  Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

    ```{r}
    #| eval: false

    ggplot(
      data = penguins,
      mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
    ) +
      geom_point() +
      geom_smooth(se = FALSE)
    ```

10. Will these two graphs look different? Why/why not? same

    ```{r}
    #| eval: false

    ggplot(
      data = penguins,
      mapping = aes(x = flipper_length_mm, y = body_mass_g)
    ) +
      geom_point() +
      geom_smooth()

    ggplot() +
      geom_point(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g)
      ) +
      geom_smooth(
        data = penguins,
        mapping = aes(x = flipper_length_mm, y = body_mass_g)
      )
    ```

# **1.4.3 Exercises**
