---
title: "Exploratory Data Analysis"
output: html_notebook
---

# Exploratory Data Analysis

-   Generate questions about your data.
-   Search for answers by visualizing, transforming, and modelling your data.

Use what you learn to refine your questions and/or generate new questions.

```{r}
library(tidyverse)
```

## Variation

-   What type of variation occurs within my variables?
-   What type of covariation occurs between my variables?

```{r}
ggplot(diamonds, aes(carat)) +
  geom_histogram(binwidth = 0.5)
```

### Typical Values

-   Which values are the most common? Why?
-   Which values are rare? Why? Does that match your expectations?
-   Can you see any unusual patterns? What might explain them?

```{r}
smaller <- diamonds |> 
  filter(carat < 3)

ggplot(smaller, aes(carat)) +
  geom_histogram(binwidth = 0.01)
```

-   Why are there more diamonds at whole carats and common fractions of carats?
-   Why are there more diamonds slightly to the right of each peak than there are slightly to the left of each peak?

Visualizations can also reveal clusters, which suggest that subgroups exist in your data. To understand the subgroups, ask:

-   How are the observations within each subgroup similar to each other?

-   How are the observations in separate clusters different from each other?

-   How can you explain or describe the clusters?

-   Why might the appearance of clusters be misleading?

### Unusual values

Outliers

```{r}
diamonds |> 
  ggplot(aes(y)) +
  geom_histogram(binwidth = 0.5)
```

The large range of the x-axis shows that there are outliers.

Change the y-coordinates to magnify the outliers.

```{r}
diamonds |> 
  ggplot(aes(y)) +
  geom_histogram(binwidth = 0.5) +
  coord_cartesian(ylim = c(0,50))
```

```{r}
unusual <- diamonds |> 
  filter(y < 3 | y > 20) |> 
  select(price, x, y, z) |> 
  arrange(y)
unusual
```
